---
title: "ProblemSet5"
author: "Yangning Tan"
format: html
editor: visual
---

## Problem 1 - Plotting

We first import the data set.

```{r}
nnmaps <- read.csv("/Users/tyn/Documents/R/data set/chicago-nmmaps.csv")
```

a.  Produce a graph of the **mean** monthly temperature in **celsius**. Draw a scatter-plot where the x-axis is month and the y-axis is the average monthly temperature in celsius. Add a line connecting the points within each season and color the lines and points by season (same color for lines and points in the same season). Be sure both x-axis and the legend are ordered appropriately, and that all labels/titles are clear and ready for publication (e.g. no variable names).

    ```{r}
    library(ggplot2)
    library(tidyverse)
    # Convert Fahrenheit to Celsius
    nnmaps$temp <- (nnmaps$temp - 32) * 5/9

    # Calculate mean monthly temperature in Celsius
    monthly_mean_temp <- nnmaps %>%
      group_by(season, month) %>%
      summarize(mean_temp = mean(temp))

    # Set plot theme
    #theme_set(theme_minimal())

    # Plot the scatter plot with lines
    ggplot(monthly_mean_temp, aes(x = month, y = mean_temp, color = season, group = season)) +
      geom_point() +
      geom_line() +
      labs(
        title = "Mean Monthly Temperature in Celsius by Season",
        x = "Month",
        y = "Mean Monthly Temperature (°C)"
      ) +
      theme(legend.position="top") +
      scale_color_manual(values = c('Winter' = 'blue', 'Spring' = 'green', 'Summer' = 'red', 'Fall' = 'orange'))
    ```

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
